---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# cuscirrigationsurvey

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

_cuscirrigationsurvey_ is an *R* package that contains survey data compiled from South Carolina County Extension Agents of Clemson University annually from 1997-2000. Some of this data was published previously in Irrigation Journal (2001), but _cuscirrigationsurvey_ is likely the most thorough compilation of that work available today. The package author received the data from W. Bryan Smith, Area Extension Agent and director of the irrigation survey efforts from 1998-2000. The survey was conducted prior to 1997, however that data has not been found or made available to the package author.

The data was gathered each year by county agents for their own county, most of whom either sat down with the growers or knew their irrigated acreage from dealing with them. 

This dataset provides a unique perspective on irrigation in South Carolina, that is still relevant today. This survey effort was the most comprehensive ever before conducted, and there were no other comparable data collection efforts for years afterwards. Lacking comprehensive data, later reports regarding irrigation in SC estimated general declines from 2000 to 2010. More recently (~2015 - ~2020), reported irrigation in South Carolina has increased substantially. Hypothetically, some of the recent increase can be attributed to under-estimated values after the year 2000.

That's the broader context, but let's see what we can learn from this dataset specifically.

To acccess the dataset, you can install the package using R:
```{r, eval = FALSE} 
# install.package('devtools') ## install devtools if you don't have it already.
devtools::install_github("capellett/cuscirrigationsurvey")
```

Or, just scroll down and copy from the data tables below.

## Irrigation Distribution System Categories

Here's a plot of the amount of acreage irrigated by the different categories of distribution system:
  
```{r echo=F}
cuscirrigationsurvey::categories |>
  dplyr::filter(is.na(Subcategory) | Subcategory %in% c(
    'Total Sprinkler Acreage', 'Total Drip Acreage', 'Total Gravity Acreage')) |>
  dplyr::group_by(Year, Category) |>
  dplyr::summarise(`Acres (thousands)`= sum(Acres)/1000, .groups='keep') |>
  ggplot2::ggplot(
    ggplot2::aes(x=Year, y=`Acres (thousands)`, color=Category)) +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ggtitle('Irrigated Acreage by Distribution System') +
  ggplot2::theme(legend.position='bottom', legend.title=ggplot2::element_blank())
# ggplot2::ggsave('CU Irrigated Acreage by Distribution System')

## use linetype instead of color to distinguish the lines.
## gravity = solid, low flow = dashed, sprinkler = dotted, total = solid and size 1.
```
We  see that sprinkler irrigation is the largest and fastest growing kind of irrigation distribution system, accounting for a majority of total irrigated acreage and nearly all of the increase in irrigated acreage from 1998-2000. Now let's look closer at each of these three categories of irrigation distribution system.

```{r echo=F}
## then each distribution system category gets a statewide graph of subcategories.
cu_irrigation_subcategory <- cuscirrigationsurvey::categories |> dplyr::filter(
  Category %in% c("Sprinkler", "Gravity/Flood", "Low Flow / Drip") &
    !(stringr::str_sub(Subcategory, 0,1) == '%')) |>
  dplyr::mutate(Acres = dplyr::if_else(is.na(Acres), 0, Acres)) |>
  dplyr::group_by(Year, Category, Subcategory) |>
  dplyr::summarise(Acres = sum(Acres, na.rm=T), .groups='keep') |>
  dplyr::ungroup()

cu_irrigation_subcategory |>
  dplyr::group_by(Category) |>
  dplyr::group_walk(function(.x, .y) {
    p <- ggplot2::ggplot(.x, ggplot2::aes(x=Year, y=Acres, color=Subcategory)) +
      ggplot2::geom_line() +
      ggplot2::theme_bw() +
      ggplot2::ggtitle(paste0('Acreage Irrigated by ', .y$Category)) +
      ggplot2::theme(legend.title=ggplot2::element_blank())
    print(p)
  })
```
Here we see that Center Pivot/Linear Move sprinklers are the most common kind of sprinkler; surface drip irrigation is the most common kind of Low Flow / Drip irrigation; and Subsurface / Seepage is the most common kind of Flood / Gravity irrigation. There are many other kinds of irrigation distribution systems in South Carolina, but they were not common during this period (and generally still aren't today).

```{r echo=F}
cuscirrigationsurvey::categories |>
  tidyr::pivot_wider(names_from=Year, values_from='Acres') |>
  scutils::reporTable()

## and a table of major counties for each category, including the %columns for that category.
## and a table of major counties, including power unit types, wells, and pumps.
## represent the Percent data with annual bar chart

```

## Irrigated Crops

There are 44 different crops included in this dataset. Below is a plot of the top 10 irrigated crops (by acreage).
```{r echo=F}

crops2 <- cuscirrigationsurvey::crops |>
  dplyr::group_by(Year, Crop) |>
  dplyr::summarise(Acres = sum(Acres, na.rm=T), .groups='drop')

crops_rank <- crops2 |>
  dplyr::group_by(Crop) |>
  dplyr::summarise(Acres=sum(Acres, na.rm=T)) |>
  dplyr::mutate(crop_rank = rank(-Acres)) |>
  dplyr::select(Crop, crop_rank) |>
  dplyr::arrange(crop_rank)

crops3 <- crops2 |>
  dplyr::left_join(crops_rank, 'Crop')
```

```{r echo=F}
crops3 |>
  dplyr::filter(crop_rank < 11) |>
  ggplot2::ggplot(
    ggplot2::aes(x=Year, y=Acres, group=Crop, color=Crop)) +
  ggplot2::geom_line() +
  ggplot2::theme_bw()


```

Corn and cotton were the two major irrigated crops, and there are many more minor irrigated crops. It isn't easy to see from the plot, but the third ranking irrigated crop is "Tree Fruits," which really means peaches mostly. 

```{r echo=F}
cuscirrigationsurvey::crops |>
  tidyr::pivot_wider(names_from=Year, values_from='Acres') |>
  scutils::reporTable()

## and a table of major counties for each category, including the %columns for that category.
## and a table of major counties, including power unit types, wells, and pumps.
## represent the Percent data with annual bar chart

```

## Irrigation system  attributes

There is still more information in this survey, which I've put in to a separate table. This includes information regarding the irrigation power sources, wells, and pumps. 

```{r echo=F} 
cuscirrigationsurvey::categories |>
  tidyr::pivot_wider(names_from=Year, values_from='Acres') |>
  scutils::reporTable()
```


```{r eval=F}
## This information is tabulated as the percentages in each county. 
## To estimate state-wide percentages, I'll weight the county percentages 
## by the irrigated area in each county.

total_acreage <- categories |>
  dplyr::filter(is.na(Subcategory)) |>
  dplyr::select(County, Year, Acres)

irr_attributes_statewide <- irr_attributes |>
  dplyr::filter(!(Subcategory %in% c('Ave Water Depth', 'Ave. Depth', 'Number'))) |>
  dplyr::left_join(total_acreage, by=c("County", "Year")) |>
  dplyr::mutate(Value = Value*Acres) |>
  dplyr::select(County, Category, Subcategory, Year, Value) |>
  dplyr::group_by(Category, Subcategory, Year) |>
  dplyr::summarise(Value=sum(Value, na.rm=T), .groups='drop')

irr_attributes |>
  dplyr::select(Category, Subcategory) |>
  unique()
```







```{r pressure, echo = FALSE}
## You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
## You can also embed plots, for example:
# plot(pressure)
```


