---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# cuscirrigationsurvey

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

_cuscirrigationsurvey_ is an *R* package that contains survey data compiled from South Carolina County Extension Agents of Clemson University annually from 1997-2000. Some of this data was published previously in Irrigation Journal (2001), but _cuscirrigationsurvey_ is likely the most thorough compilation of that work available today. The package author received the data from W. Bryan Smith, Area Extension Agent and director of the irrigation survey efforts from 1998-2000. The survey was conducted prior to 1997, however that data has not been found or made available to the package author.

The data was gathered each year by county agents for their own county, most of whom either sat down with the growers or knew their irrigated acreage from dealing with them. 

This dataset provides a unique perspective on irrigation in South Carolina, that is still relevant today. This survey effort was the most comprehensive ever before conducted, and there were no other comparable data collection efforts for years afterwards. Lacking comprehensive data, later reports regarding irrigation in SC estimated general declines from 2000 to 2010. More recently (~2015 - ~2020), reported irrigation in South Carolina has increased substantially. Hypothetically, some of the recent increase can be attributed to under-estimated values after the year 2000.

To acccess the dataset, you can install the package using R:
```{r, eval = FALSE} 
# install.package('devtools') ## install devtools if you don't have it already.
devtools::install_github("capellett/cuscirrigationsurvey")
```

Or, just check out the article to get started.



